---
title: "YFRW Modelling"
output: html_notebook
---

# Setup - load files, colours, data management

```{r setup}
library(pacman)
p_load(readxl, tidyverse, mgcv, mgcViz, car, RColorBrewer, wesanderson, sf, terra, mapview)
setwd("C:/Users/laure/OneDrive/Desktop/PhD/Modelling/New GAM")
Olary_new <- readxl::read_excel("../../../data/final/Olary_new.xlsx")
Flinders_new <- readxl::read_excel("../../../data/final/Flinders_new.xlsx")
ALA_ALL <- readxl::read_excel("../../../data/final/ALA_ALL.xlsx")
all_raw <- readxl::read_excel("../../../data/final/Combined_survey.xlsx")
# all_30 is a derivation of all_raw with counts averaged for the geometry of jaxa dem grid cells, Australian Lambert (3112)
all_30 <- read.csv("../../../data/final/all_obsBO_30_ExportTable.txt")
```


```{r colours}
# choose distinct colours -  wesanderson palettes
#  https://github.com/karthik/wesanderson
names(wes_palettes)
mycols_wes <- c(wes_palette("Darjeeling2"), wes_palette("Darjeeling1"),
                wes_palette("Chevalier1"), wes_palette("BottleRocket2"), 
                wes_palette("Moonrise1"), wes_palette("Moonrise2"), 
                wes_palette("Moonrise3"), wes_palette("GrandBudapest1"), 
                wes_palette("GrandBudapest2"), wes_palette("IsleofDogs1") 
                )
mycols_wes <- mycols_wes[-5]
scales::show_col(mycols_wes)
scales::show_col(mycols_wes[1:5])

```

```{r data prep Presence count5 Region}
# rename grid_code to count
names(all_30)[2] <- "count"
# add binary presence/absence
all_raw$Presence <- as.factor(all_raw$count > 0)
all_30$Presence <- as.factor(all_30$count > 0)
# add one to counts to allow log transform
all_raw$count1 <- all_raw$count + 1
all_30$count1 <- all_30$count + 1
# classes
all_raw$count5 <- floor(all_raw$count/5)
all_raw$count5 <- ifelse(all_raw$count5 >= 5, 5, all_raw$count5)
all_30$count5 <- floor(all_30$count/5)
all_30$count5 <- ifelse(all_30$count5 >= 5, 5, all_30$count5)

all_30 <- all_30 %>%
  mutate(count5 = floor(count/5)) %>%
  mutate(count1 = count + 1,
         count5 = ifelse(count5 >= 5, 5, count5),
         Region = ifelse(X > 545000, "Olary_Ranges", "Flinders_Gammon_Ranges"))
 
hist(all_30$count)
hist(all_30$count1)
hist(all_30$count5)
hist(all_raw$count)
hist(all_raw$count5)
hist(all_raw$count1)

```


```{r data prep groups of locations}
# Convert table to features
points30 <- st_as_sf(all_30, coords = c("X", "Y"), crs = 3112, remove=F)
# generate projected layer from lat/lon to ensure it's Geoscience Lambert Australia
points <- st_as_sf(all_raw, coords = c("lon", "lat"), crs = 4326, remove=F) %>% 
  st_transform(3112) %>%
  # update X and Y
  select(-c("X", "Y")) %>% 
  cbind(st_coordinates(.))

# Create spatially disjoint groups of points
# buffer to generate reasonably large clusters of points 10000 gives 5 
xx <- st_buffer(points, dist = 10000, nQuadSegs = 4) %>%
  st_union() %>%  # union to combine overlapping buffers
  st_cast("POLYGON")  # convert multipolygon to single polygons

# find groupID (polygon ID at each point)
groupID <- unlist(st_intersects(points, xx))

# add groupID to points
points$groupID <- (as.factor(groupID))

# repeat for all_30
xx <- st_buffer(points30, dist = 10000, nQuadSegs = 4) %>%
  st_union() %>%  # union to combine overlapping buffers
  st_cast("POLYGON")  # convert multipolygon to single polygons

# find groupID (polygon ID at each point)
groupID <- unlist(st_intersects(points30, xx))

# add groupID to points30
points30$groupID <- (as.factor(groupID))

# find number of locations per group
sites <- points %>%
  group_by(groupID) %>% 
      summarise(n = n()) %>%
  st_drop_geometry()
sites

# find number of locations per group for points30
sites30 <- points30 %>%
  group_by(groupID) %>% 
      summarise(n = n()) %>%
  st_drop_geometry()
sites30
rm(xx)
gc()
```

```{r data prep nice names}
names(points)
#  [1] "LOCATIONID" "solar_sum"  "solar_wint" "ruggedness" "dem"        "OBJECTID"   "id"        
#  [8] "lon"        "lat"        "time"       "Region"     "count"      "orgnl_d"    "observr"   
# [15] "notes"      "slope"      "aspect"     "Presence"   "count1"     "count5"     "X"         
# [22] "Y"          "geometry"   "groupID"   
nice_names_raw <- tibble(original_names = names(points), 
                         txt = c("LOCATIONID", "Solar Radiation, Summer", "Solar Radiation, Winter",  "Terrain Ruggedness", "Elevation",  "OBJECTID",   "id", "Longitude",   "Latitude", "time",  "Region",  "Scat count (m-2)", "orgnl_d",    "observr",    "notes",   "Terrain Slope", "Terrain Aspect","Scat Presence", "Scat count + 1", "Scat class", "x", "y", "geometry", "Group ID"))
nice_names_raw

names(points30)
#  [1] "pointid"           "count"         "X"                 "Y"                
#  [5] "directSolarSummer" "directSolarWinter" "rugged_clip"       "jaxa_dem"         
#  [9] "NEW_jaxa_clip"     "Slope_NEW_DEM"     "Aspect_NEW_DEM"    "Current_Bio_1"    
# [13] "Current_Bio12"     "Current_YFRW_SDM"  "Presence"          "count1"           
# [17] "count5"            "Region"            "geometry"          "groupID"          
nice_names <- tibble(original_names = names(points30),
                       names = c("pointid", "count", "X", "Y", "solar_sum", "solar_wint", "ruggedness",  "jaxa_dem", "dem",  "slope",  "aspect", "T", "rain", "YFRW_SDM", "Presence", "count1", "count5", "Region", "geometry", "groupID"),
                       txt = c("pointid", "Scat count (m-2)", "x", "y", 
  "Solar Radiation, Summer", "Solar Radiation, Winter",  "Terrain Ruggedness",  "jaxa_dem", 
  "Elevation",  "Terrain Slope",  "Terrain Aspect", "Temperature", "Rainfall", 
  "YFRW SDM, Current Climate", "Scat Presence", "Scat count + 1", "Scat class", "Region", "geometry", "Group ID"))
names(points30) <- nice_names$names
nice_names

write_sf(points, "../../final/all_obs.shp", overwrite = TRUE)
write_sf(points30, "../../final/all_30.shp", overwrite = TRUE)

# create nice colours for Flinders and Olary
if(nrow(sites30) == 5) {
  site_colours <- mycols_wes[c(2,12,5,6,14)]
} else {
  # automatic for variable numbers
  fg_sites <- points %>% 
    st_drop_geometry() %>%
    filter(Region == "Flinders_Gammon_Ranges") %>% 
    select(groupID) %>% pull() %>%
    unique() 
  ol_sites <- points %>% 
    st_drop_geometry() %>%
    filter(Region == "Olary_Ranges") %>% 
    select(groupID) %>% pull %>%
    unique()
  flinders_cols <- mycols_wes[c(3:(2 + length(fg_sites)))]
  olary_cols <- mycols_wes[c(12:(11 + length(ol_sites)))]
  # create colours for sites
  site_colours <- c(flinders_cols, olary_cols)
}
scales::show_col(site_colours)

mapview(points, zcol = "groupID", alpha = .2, col.regions = site_colours)
mapview(points30, zcol = "groupID", alpha = .2, col.regions = site_colours)
ggsave("points30_groups.jpg", width = 15, height = 12, units = "cm", dpi = 1200 )

# create a copy with a short name and geometry dropped
all <- points30 %>%
  st_drop_geometry()
all$Region <- as.factor(all$Region)





# First series of model testing - single factors

```{r series 1 - dem}
F0D <- gam(log(count1) ~ s(dem) + Region , data = all,  method = "REML", family = tw())
summary(F0D)
FP <- getViz(F0D)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)

```

```{r series 1 - ruggedness}
F0R <- gam(log(count1) ~ s(ruggedness) + Region , data = all,  method = "REML", family = tw())
summary(F0R)
FP <- getViz(F0R)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)

```
```{r series 1 - slope}
F0S <- gam(log(count1) ~ s(slope) + Region , data = all,  method = "REML", family = tw())
summary(F0S)
FP <- getViz(F0S)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)

```

```{r series 1 - aspect}
F0A <- gam(log(count1) ~ s(aspect) + Region , data = all,  method = "REML", family = tw())
summary(F0A)
FP <- getViz(F0A)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)

```

```{r series 1 - solar wint}
F0SW <- gam(log(count1) ~ s(solar_wint) + Region , data = all,  method = "REML", family = tw())
summary(F0SW)
FP <- getViz(F0SW)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)

```

```{r series 1 - solar sum}
F0SS <- gam(log(count1) ~ s(solar_sum) + Region , data = all,  method = "REML", family = tw())
summary(F0SS)
FP <- getViz(F0SS)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)

```

# Second series of model testing - multifactorial 

```{r series 2 - Base model}
FBase <- gam(log(count1) ~ s(dem) +s(ruggedness) + Region , data = all,  method = "REML", family = tw())
summary(FBase)
FP <- getViz(FBase)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)

```

```{r series 2 - Base + slope}
FDRS <- gam(log(count1) ~ s(dem) + s(ruggedness) + s(slope) + Region , data = all,  method = "REML", family = tw())
FP <- getViz(FDRS)
summary(FDRS)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)
anova.gam(FBase, FDRS, test = "F")


```{r series 2 - Base + aspect}
F4a <- gam(log(count1) ~ s(dem) +  s(ruggedness) + s(aspect) + Region, data = all,  method = "REML", family = tw())
summary(F4a)
FP <- getViz(F4a)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)
anova.gam(FBase, F4a, test = "F")
```

```{r series 2 - Base + winter solar radiation}
F4b <- gam(log(count1) ~ s(dem) +  s(ruggedness) + s(solar_wint) + Region, data = all,  method = "REML", family = tw())
summary(F4b)
FP <- getViz(F4b)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)
anova.gam(FBase, F4b, test = "F")
```

```{r series 2 - Base + summer solar radiation}
F4c <- gam(log(count1) ~ s(dem) +  s(ruggedness) + s(solar_sum) + Region, data = all,  method = "REML", family = tw())
summary(F4c)
FP <- getViz(F4c)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)
anova.gam(FBase, F4c, test = "F")
```

```{r series 2 - Base + dem/ruggedness interaction - FINAL MODEL}
F3 <- gam(log(count1) ~ s(dem) + s(ruggedness) + te(dem, ruggedness) + Region , data = all,  method = "REML", family = tw())
summary(F3)
FP <- getViz(F3)
print(plot(FP, allTerms = T), pages = 1)
gam.check(FP)
anova.gam(FBase, F3, test = "F")
```

## Visualisation of scat counts by region
```{r Figure of presence/absence using dem and ruggednes}

vars <- c(which(nice_names$names == "dem"), which(nice_names$names == "ruggedness")) 
lables <- nice_names$txt[vars]

all %>%
  filter(Region != "Flinders_Gammon_Ranges") %>%
  mutate(Presence = as.factor(count > 0)) %>%
  ggplot(aes(x=dem, y=ruggedness, size = Presence, shape = Presence, colour = Presence)) +
  geom_point(alpha = 0.5) + 
  scale_color_manual(values=c('#E69F00', 'darkblue')) +
  scale_shape_manual(values=c(16, 3))+
  scale_size_manual(values=c(.5, 1))+
  labs(x = lables[1], y = lables[2], title = "Olary Ranges") +
  theme(legend.position="bottom") + 
  theme_minimal()
ggsave("dem_rug_OL.jpg", width = 15, height = 10, units = "cm", dpi = 1200 )

all %>%
  filter(Region == "Flinders_Gammon_Ranges") %>%
  mutate(Presence = as.factor(count > 0)) %>%
  ggplot(aes(x=dem, y=ruggedness, size = Presence, shape = Presence, colour = Presence)) +
  geom_point(alpha = 0.5) + 
  scale_color_manual(values=c('#E69F00', 'darkblue')) +
  scale_shape_manual(values=c(16, 3))+
  scale_size_manual(values=c(.5, 1))+
  labs(x = lables[1], y = lables[2], title = "Flinders/Gammon Ranges") +
  theme(legend.position="bottom") + 
  theme_minimal()
ggsave("dem_rug_FG.jpg", width = 15, height = 10, units = "cm", dpi = 1200 )

all %>%
  mutate(Presence = as.factor(count > 0)) %>%
  ggplot(aes(x=dem, y=ruggedness, size = Presence, shape = Presence, colour = Presence)) +
  geom_point(alpha = 0.5) + 
  scale_color_manual(values=c('#E69F00', 'darkblue')) +
  scale_shape_manual(values=c(16, 3))+
  scale_size_manual(values=c(.5, 1))+
  labs(x = lables[1], y = lables[2], title = "Entire area") +
  theme(legend.position="bottom") + 
  theme_minimal()
ggsave("dem_rug_all.jpg", width = 15, height = 10, units = "cm", dpi = 1200 )

```


##Appendix Figures
## Correlation of variables

```{r Appendix Figure S7- correlation matrix by GroupID}
# remove sites with low number of members and take sample for better visualisation

all_select <- points30 %>%
  st_drop_geometry() %>%
  sample_n(1000) %>%
  droplevels

scatterplotMatrix(~log(count1)+dem+ruggedness+slope+aspect+solar_sum+solar_wint|groupID, 
                  data=all_select,
                  col=site_colours,
                  regLine=F,
                  smooth = list(method=gamLine, k=30),
                  cex=0.2,
                  main="Scatter plot by spatial group - 30m resampled"
  )

```

```{r Appendix Figure S9- dem and ruggedness for all five groups}
vars <- c(which(nice_names$names == "dem"), which(nice_names$names == "ruggedness")) 
lables <- nice_names$txt[vars]

all %>%
  ggplot(aes(x=dem, y=ruggedness, shape = Presence, size = log(count1), colour = groupID)) +
  geom_point(alpha = .5) + 
  scale_color_manual(values=site_colours) +
  scale_shape_manual(values=c(16, 3))+
  scale_size_continuous(range = c(0.1, 2)) +
  labs(x = lables[1], y = lables[2], title = "Entire region") +
  theme(legend.position="bottom") +   theme_minimal()
ggsave("dem_rug_allgroups.jpg", width = 15, height = 10, units = "cm", dpi = 1200 )

```

